Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-21 Manuscript/manuscript.Rmd
improving story

We then investigated which genomic annotations were most predictive of TAD boundary formation, across resolution for random forest models with the best given set of data-level characteristics determined above (SMOTE re-sampled with distance type predictors). We found that specific transcription factor binding sites (TFBS), including SMC3ab, ZNF143, CTCF, and RAD21, were given consistently high rankings of predictive importance in the models built on the GM12878 cell line (Table 1). These are known factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] and were found to be ranked consistently high on the K562 cell line as well, with similar rankings across resolutions (Supplementary Table 6). Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1  transcription factor, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation histone modifications were ranked high on the 10 kb resolution, but were less consistently ranked at lower resolutions for both cell lines. <!--Moreover, the ranking profiles of genomic annotations for both cell lines were found to be more similar between the 10 kb and 25 kb resolutions, as well as between the 50 kb and 100 kb resoluions (this is proved when looking at a clustered heatmap).-->

>>>>

In addition to adressing both the affect of class imbalance and the optimal spatial relationships used when predicting TAD boundaries, we also aimed at identifying the genomic annotations that were most predictive of TAD boundary formation. In doing so, we investigated which genomic annotations yielded the hight predictive importance values for random forest models implemented on the best given set of data-level characteristics determined above (SMOTE re-sampled with distance type predictors), across resolution and cell line. We found that specific transcription factor binding sites (TFBS), including SMC3ab, ZNF143, CTCF, and RAD21, were given consistently high predictive importance in the models built on the GM12878 cell line, across all resolutions (Figure 8). These are known factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] and were found to yield consistently high predictive importance on the K562 cell line as well, with similar profiles across resolutions (Supplementary Figure 15). Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1  transcription factor, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation histone modifications were ranked high on the 10 kb resolution, but were less consistently ranked at lower resolutions for both cell lines. <!--Moreover, the ranking profiles of genomic annotations for both cell lines were found to be more similar between the 10 kb and 25 kb resolutions, as well as between the 50 kb and 100 kb resoluions (this is proved when looking at a clustered heatmap).-->
